Pleiotropy of gene regulatory variants increases with the number of regulated genes, active cell types and tissues and bound transcription factors

### 2022-11-12 ms/ms01_titlepage.tex
ENH ms and refactoring
ENH updated with 418
MS

\title{Trait pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
%\shorttitle{Pleiotropy and gene regulatory variants}

@@

Pleiotropic genetic variants affecting several traits are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
%
Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
%
In this work, I investigate the link between gene regulation and trait pleiotropy for frequent variants.

% Verify numbers
First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
%
I have categorized the GWAS traits into 96 traits and categories to define pleiotropic as belonging at least to two of these categories.
%
Pleiotropic variants are associated to more eQTL genes (eGenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
%
These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
%
These variants with the GWAS traits, egenes and etissues can be explored in a public website.

@@

\begin{keywords2}
eGenes; eQTLs
\end{keywords2}

>>>>

\title{Pleiotropy of gene regulatory variants increases with the number of regulated genes, active cell types and tissues and bound transcription factors}
%\shorttitle{Pleiotropy and gene regulatory variants}

@@

Pleiotropic genetic variants affecting several traits are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
%
Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
%
In this work, I investigate the link between frequent gene regulatory variants and pleiotropy.

% Verify numbers
First I have computed the colocalization probability of 418 GWAS and 127 eQTL studies and I have observed 10,334 GWAS/eQTL pairs with at least one colocalized variants with a probability above 0.8.
%SELECT count(DISTINCT concat(`gwas_id`, `eqtl_id`)) FROM `gwas2eqtl_modelgwas2eqtl` where `PP_H4_abf`>=0.8; 10334
%Pourcentage of GWAS
%
I have categorized the GWAS traits into 96 traits and categories to define pleiotropic as belonging at least to two of these categories.
%
Pleiotropic variants are associated to more eQTL genes (eGenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
%
These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
%
These variants with the GWAS traits, egenes and etissues can be explored in a public website.

@@

% Keywords
% eGenes; eQTLs
